---
title: Financial markets dashboard
##author: |
##  | Professor Robert McDonald
##  | Kellogg School, Northwestern University
##date: "`r format(Sys.time(), '%A, %B %d, %Y %H:%M %Z')`"
date: now
format: 
    html:
        embed-resources: true
---


```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(plotly)

theme_set(theme_bw())

getdatanomatterwhat <- TRUE
getdatanomatterwhat <- FALSE
filepath <- 'data/'
fn <- 'dashboard_data.Rdata'
datafile <- paste0(filepath,fn)

## DATA UPDATES
## Can set an age check on data file to determine if re-read is needed
updateinterval <- 0.9 ## days after which the data will be updated
if (file.exists(datafile)) {
    filetime <- file.mtime(datafile)
    time_since_update <-
        (as.integer(Sys.time()) - as.integer(filetime))/(3600*24)
    getdata <- ifelse(time_since_update < updateinterval, FALSE, TRUE)
} else {
    getdata <- TRUE
}

if (getdatanomatterwhat) getdata <- TRUE  ## ignore updateinterval

## DATA CHARACTERISTICS
startdate <- '2023-01-01'

library(knitr)
opts_chunk$set(echo = FALSE
              ,message = FALSE
              ,warning = FALSE
               )
options(scipen = 100) ## no scientific notation
```

```{r getdata}
library(tidyquant)
stocks <- c('^GSPC', '^VIX', '^SP600', 'BTC-USD')
currs <- c('DEXUSEU', 'DEXUSUK', 'DEXJPUS')
treas <- c('1-year' = 'DGS1', '10-year' = 'DGS10', '5-year' = 'DGS5',
           '3-month' = 'DTB3')

if (getdata) {
    x <- tq_get(stocks, from = startdate) %>%
        select(symbol, date, price = adjusted) %>%
        mutate(symbol = gsub('\\^GSPC', 'SP500', symbol),
               symbol = gsub('\\^SP600', 'SP600', symbol),
               symbol = gsub('\\^VIX', 'VIX',  symbol))
    xex <- tq_get(currs, from = startdate, get = 'economic.data') %>%
        mutate(symbol = gsub('DEX', '', symbol))
    rates <- tq_get(unname(treas), from = startdate, get = 'economic.data' ) %>%
        rename(rate = price)
    for (i in names(treas)) rates$symbol <- gsub(paste0('^', treas[i], '$'), i,  rates$symbol)
    save(x, xex, rates, file = datafile)
} else {
    load(datafile)
}
filetime <- file.mtime(datafile)
```

```{r functions}
sumtbl_old <- function(df, volret = TRUE, symorder = NULL) {
    if (is.null(symorder)) symorder <- unique(df$symbol)
    names(df)[3] <- 'price'
    out <- list()
    for (i in symorder) {
        out[[i]] <- filter(df, symbol == i) %>%
            na.omit() %>% 
            summarize(Symbol = i,
                      Min = min(price, na.rm = TRUE),
                      Max = max(price, na.rm = TRUE),
                      Initial = first(price),
                      Last =  last(price))
    }
    return(do.call(rbind, out))
}

sumtbl <- function(df, volret = TRUE, symorder = NULL) {
    names(df)[3] <- 'price'
    ##df <- x
    out <- df %>% 
        na.omit() %>%
        group_by(symbol) %>% 
        summarize(Min = min(price, na.rm = TRUE),
                  Max = max(price, na.rm = TRUE),
                  Initial = first(price),
                  Last =  last(price)) 
    return(out)
}


```

# Select market data {.tabset .tabset-pills}

Date and time of last data update: `r format(filetime, "%A %B %d %Y %H:%M %Z")`

::: {.panel-tabset}

## Stock market and bitcoin


```{r plotstocks, eval=TRUE}

## Compute table with min, max, std dev, last week, last month, sample
tickerorder <- c('SP500', 'SP600', 'VIX', 'BTC-USD')
x$symbol <- factor(x$symbol, levels = tickerorder)
p = ggplot(x, aes(x = date, y = price)) +
    geom_line() +
    scale_y_log10() + #limits = c(10, NA)) +
    facet_wrap(~ symbol, scales = 'free_y', ncol = 2
               ) +
    labs(caption = "Source: Yahoo, via R's tidyquant package") + 
    ggtitle('Stock and crypto market data, log scale')

ggplotly(p, tooltip = c('date', 'price'))
kable(sumtbl(x, symorder = tickerorder))
```
Note: The S&P 600 index is S&P's "small-cap" index, containing firms
with market capitalizations roughly between $1 billion and $4 billion.
The S&P 400 is the "mid-cap" index, and the S&P 1000 is the combination of the two.

## Foreign Exchange

```{r plotfx}
p <- ggplot(xex, aes(x = date, y = price)) +
    geom_line() +
    scale_y_log10() +
    facet_wrap(~ symbol, scales = 'free_y', ncol = 1) +
    ggtitle('Exchange rates, log scale' )


ggplotly(p, tooltip = c('date', 'price'))
kable(sumtbl(xex))
```


## Interest rates

```{r plotrates, results='asis'}
rateorder <- c('3-month','1-year','5-year','10-year')
rates$symbol <- factor(rates$symbol, levels = rateorder)
p <- ggplot(rates, aes(x = date, y = rate)) +
    geom_line() +
    facet_wrap(~ symbol, ncol = 2) +
    ggtitle('Treasury rates, various maturities')

ggplotly(p, tooltip = c('date', 'rate'))
kable(sumtbl(rates, symorder = rateorder))
```


## Data sources and disclaimer


### Data sources

All data was retrieved using R's [tidyquant
package](https://cran.r-project.org/package=tidyquant). The
underlying data sources were:

* **Stock data** [Yahoo Finance](https://finance.yahoo.com/) 
  * Symbols: `r stocks`

* **Exchange rate data** [St Louis Federal Reserve FRED data
  repository](https://fred.stlouisfed.org)
  * Symbols: `r currs`

* **Interest rate data** [St Louis Federal Reserve FRED data
  repository](https://fred.stlouisfed.org)
  * Symbols: `r unname(treas)`
  
  
### Disclaimer

This site was created for **educational purposes**. There is **no warranty**
as to the accuracy or timeliness of the data. If you're tempted to
base trading decisions on this data, you should 

1. Take several slow, deep breaths
1. Do your own independent research
2. Take an investments course, or at least read an investments book ---
   [Bodie, Kane, and
   Marcus](https://www.mheducation.com/highered/product/investments-bodie-kane/M9781260013832.html)
   is excellent

### Purpose

This is a simple document. I wrote it to provide a focal point for
   class discussions, focusing on information that I hope will be
   salient for students. Comments are welcome. It is licensed under
   the MIT License and [the code is available on
   Github](https://github.com/rmcd1024/fin-dash)
   
-- [Robert McDonald](https://www.kellogg.northwestern.edu/faculty/directory/mcdonald_robert.aspx), [Kellogg
School of Management](https://kellogg.northwestern.edu), Northwestern University

:::
